Forecasting Multi-District Elections

### 2019-11-25 outputs/paper/paper.Rmd
just some edits to tighten up the script a little. All the diffs in the data are just bringing my branch into the master (up to date stats).

## Reasonable estimates of electoral division counts
The first aspect of our framework is that a useful forecasting model of multi-district elections provides probabilistic output from an interpretable model that is ex anti consistent with the available evidence and ex post only inaccurate for discernable reasons.

A reasonable model must be interpretable because it is unavoidably partially assessed on its explanatory value, as elections occur only every few years. @dowding2019prediction [p. 1,001] distinguish between 'pragmatic prediction' which is what we have engaged in here, and 'scientific prediction', and argue that '[a]ssessing how good [pragmatic] predictions are is based on how closely the assigned probabilities match what actually happens.' But given how irregular and idiosyncratic each election is, a useful forecasting model would also engage in 'scientific prediction', in which the explanatory value of the model is also important. Although substantial improvements to a useful forecasting model of multi-district elections would occur after each election, an interpretable model allows continuous smaller-scale improvements. Additionally, one reason that political opinion polling occurs is so that campaigns and the media can use it. An uninterpretable model would be limiting in this regard.

@@

## Account for preferences or turnout

@@

Preferences are crucial in Australian elections, and much effort is spent when conducting political opinion polling to decide how best to distribute them for their estimates. It is easy to do this in a deterministic way, for instance, by assuming that what happened last election happens again, or to assume some other reasonable sounding but similar approach. We argue that a useful forecasting model would account for preferences in a probabilistic way.

@@

Reproducibility refers to the ability of a third-party to repeat a process and get essentially the same outcome. This is the minimum that is required of credible modern quantitative approaches, and in its absence credibility requires 'trust in the confidence or authority of the originator' [@munafo2017manifesto, p. 5]. Given how easy it is for the incentives of those forecasting multi-district elections to be questioned we contend reproducibility also ought to be the minimum required of a useful forecasting model for multi-district elections. But given general practise, we would suggest that greater transparency is needed before reproducibility is possible. To a certain extent, recommendations from industry associations such as the American Association for Public Opinion Research (AAORP) or, in Australia, the Association of Market and Social Research Organisations (AMSRO) are also concerned with this.

@jamieson2019signaling [p. 19,233] describe transparency '[a]s a prerequisite to scrutiny' and argue that it allows 'others to examine study design, execution, and data and, as a result, replicate results.' However, we go further than @jamieson2019signaling and argue that a useful forecasting model would not just require '[a]rchiving data and analysis plans in publicly available repositories', but would also impose a timing and access constraint. Specifically, the data and analysis of a useful forecasting model of multi-district elections must be available at the same time as any public release relating to the forecasting model, be that an academic research paper or press coverage, and it must be just as easily accessed. For instance, it is not enough that datasets and analysis be available on request, if the coverage of the forecasting model is in the media; they must be able to be accessed as easily as the media, for instance in a publicly available repository.

@@

The fourth aspect of a useful forecasting model of multi-district elections is that its central results should be inextricably linked to their distribution, and these results should be communicated in a way that does not rely on an audience being statistically fluent. 
Although political opinion polls regularly include some measure of confidence, this tends to be separable from the central estimate, and in practice, regularly is [@larsen2019transforming]. While this is understandable given the constraints faced by journalists [@ricketson2019like] and the difficulty that even experts have appreciating the nuances of statistical concepts [@hoekstra2018improving; @greenland2016statistical], a useful model would account not allow this to happen.

@@

Finally, the results of a useful forecasting model should not require statistical fluency to understand. Misunderstandings on the part of consumers of estimates are, to a large extent, the responsibility of those who produce estimates.

@@

The combination of the levels of the characteristics characterise the sub-groups that are used for post-stratification. In this case, there are two categories for gender, four categories for age-group, and four categories for highest education obtained, so there are thirty-two sub-groups, denoted by $c \in [1, 2, \ldots, C]$, where $C=32$. The number of people in each sub-group for each electoral division is similarly denoted $N_c$. We discard informal votes. In this set-up, $\mbox{2PP}_{d,p=1}$ is the Australian Labor Party's two-party-preferred share in electoral division $d$.

@@

At this stage of our approach, the model is being trained on the raw datasets, and these results are not yet post-stratified. We have more to say about this in Section \ref{discussion}, but ours is first and foremost a forecasting rather than explanatory model, nonetheless there is value in examining the results for reasonableness. The estimates are in relation to how likely a respondent is to support the Australian Labor Party. In the case of both datasets, males are expected to be less likely to support the Australian Labor Party. Monotonically, respondents in older age-groups are more likely to support the Australian Labor Party in the Smartvote Australia dataset, but there is no monotonic trend in the Life in Australia dataset.

@@

One important aspect of post-stratification is to understand how the raw survey results are being changed. We consider this on the basis of electoral division in Figure \ref{fig:rawdatacomparedwithmodelestimates}. The estimate of the first preference share for each electoral division is made by making many forecasts of the outcome, and then examining this distribution. Figure \ref{fig:rawdatacomparedwithmodelestimates}, which follows @Alexander2019, shows the mean of the distribution and 95 per cent prediction intervals (as a reminder, prediction intervals tend to be wider than confidence intervals). These can be compared with the raw data from the survey. In general, it can be seen that the raw estimate of the Australian Labor Party's share of first preferences in both surveys is being reduced by the model. Figure \ref{fig:rawdatacomparedwithmodelestimates} is on the basis of electoral divisions, however it is also possible to do a similar exercise on the basis of, say, age-group, gender, or education.

@@

The classifications used by the Australian Electoral Commission are usually used only on actual results, and so we adjust the central classification in a way that is more explicit about uncertainty. Specifically: 'safe loss' means a two-party-preferred share of less than 0.4; 'fairly safe loss' means between 0.4 and 0.44; 'marginal loss' means between 0.44 and 0.47; 'too close to call' means between 0.47 and 0.53; 'marginal win' means between 0.53 and 0.56; 'fairly safe win' means between 0.56 and 0.6; and finally, 'safe win' means a two-party-preferred share greater than 0.6. Table \ref{tab:haveagoyougetago} shows this disaggregation.

@@

Using this classification, it can be seen that estimates based on the Life in Australia dataset tend to be more accurate than those based on the Smartvote Australia dataset.

@@

In our model we account for preferences in a way that propagates uncertainty by applying a model that relates first-preferences shares to two-party-preferred shares which has been trained on the past four elections. This results in increased variance, compared with a deterministic approach.

@@

@@

Finally, in terms of communication, we have provided some variations designed to ensure that the probability nature of our model cannot be separated from central estimates. However, this is just a first step and there is much more that could be done, including survey work and field testing to examine how this could be done in a better way.

@@

One weakness of our approach is that it is not electoral-division-specific, which means that there are nuances of specific electoral divisions that it misses. Better preferences model - take account of preferences deals, and also not lump all non-major parties in as other - separate them based on ideology so that there is at least some better flow.

>>>>

## Reasonable estimates of electoral division counts
The first aspect of our framework is that a useful forecasting model of multi-district elections provides probabilistic output from an interpretable model that is ex ante consistent with the available evidence and ex post only inaccurate for discernable reasons.

A reasonable model must be interpretable because it is unavoidably partially assessed on its explanatory value, as elections occur only every few years. @dowding2019prediction [p. 1,001] distinguish between 'pragmatic prediction' which is what we have engaged in here, and 'scientific prediction', and argue that '[a]ssessing how good [pragmatic] predictions are is based on how closely the assigned probabilities match what actually happens.' But given how irregular and idiosyncratic each election is, a useful forecasting model would also engage in 'scientific prediction', in which the explanatory value of the model is also important. Although substantial improvements to a useful forecasting model of multi-district elections would occur after each election, an interpretable model allows continuous smaller-scale improvements. <!--Additionally, one reason that political opinion polling occurs is so that campaigns and the media can use it. An uninterpretable model would be limiting in this regard. -->

@@

## Account for features of the electoral system

@@

Australia's election system follows a preferenctial voting system, in which an instant runoff election is held for each election. Estimating the distribution of preference flows from eliminated candidates is crucial to predicting the results Australian elections. It is easy to do this in a deterministic way, for instance, by assuming that what happened last election happens again, or to assume some other reasonable sounding but similar approach. We argue that a useful forecasting model would account for preferences in a probabilistic way.

@@

Reproducibility refers to the ability of a third-party to repeat a process and get essentially the same outcome. This is the minimum that is required of credible modern quantitative approaches, and in its absence credibility requires 'trust in the confidence or authority of the originator' [@munafo2017manifesto, p. 5]. Given how easy it is for the incentives of those forecasting multi-district elections to be questioned we contend reproducibility also ought to be the minimum required of a useful forecasting model for multi-district elections. But given general practise, we would suggest that greater transparency is needed before reproducibility is possible. <!--To a certain extent, recommendations from industry associations such as the American Association for Public Opinion Research (AAORP) or, in Australia, the Association of Market and Social Research Organisations (AMSRO) are also concerned with this. -->

@jamieson2019signaling [p. 19,233] describe transparency '[a]s a prerequisite to scrutiny' and argue that it allows 'others to examine study design, execution, and data and, as a result, replicate results.' However, we go further than @jamieson2019signaling and argue that a useful forecasting model would not just require '[a]rchiving data and analysis plans in publicly available repositories', but would also impose a timing and access constraint. Specifically, the data and analysis of a useful forecasting model of multi-district elections must be available at the same time as any public release relating to the forecasting model, be that an academic research paper or press coverage, and it must be just as easily accessed. <!--For instance, it is not enough that datasets and analysis be available on request, if the coverage of the forecasting model is in the media; they must be able to be accessed as easily as the media, for instance in a publicly available repository.-->

@@

The fourth aspect of a useful forecasting model of multi-district elections is that its central results should be inextricably linked to their distribution, and these results should be communicated in a way that does not rely on an audience being statistically fluent. 
Although political opinion polls regularly include some measure of confidence, this tends to be separable from the central estimate, and in practice, regularly is [@larsen2019transforming].

@@

While the tendancy to downplay uncertainty is understandable given the constraints faced by journalists [@ricketson2019like] and the difficulty that even experts have appreciating the nuances of statistical concepts [@hoekstra2018improving; @greenland2016statistical], the results of a useful forecasting model should not require statistical fluency to understand. Misunderstandings on the part of consumers of estimates are, to a large extent, the responsibility of those who produce estimates. To make the impact of uncertainty clear, communication of a model should explicitly account for areas of uncertainty in layman's terms by, for example, reporting districts that are too close to call.

@@

The combination of the levels of the characteristics characterise the sub-groups that are used for post-stratification. In this case, there are two categories for gender, four categories for age-group, and four categories for highest education obtained, so there are thirty-two sub-groups, denoted by $c \in [1, 2, \ldots, C]$, where $C=32$. The number of people in each sub-group for each electoral division is similarly denoted $N_c$. Informal votes (also known as invalid or spolit ballots in other jurisdictions) were discarded. In this set-up, $\mbox{2PP}_{d,p=1}$ is the Australian Labor Party's two-party-preferred share in electoral division $d$.

@@

At this stage of our approach, the model is trained on the raw datasets, and these results are not yet post-stratified.<!-- We have more to say about this in Section \ref{discussion} but ours is first and foremost a forecasting rather than explanatory model, nonetheless there is value in examining the results for reasonableness.--> The estimates are in relation to how likely a respondent is to support the Australian Labor Party. In the case of both datasets, males are expected to be less likely to support the Australian Labor Party. Monotonically, respondents in older age-groups are more likely to support the Australian Labor Party in the Smartvote Australia dataset, but there is no monotonic trend in the Life in Australia dataset.

@@

One important aspect of post-stratification is to understand how the raw survey results are changed. We consider this on the basis of electoral division in Figure \ref{fig:rawdatacomparedwithmodelestimates}. The estimate of the first preference share for each electoral division is made by making many forecasts of the outcome, and then examining this distribution. Figure \ref{fig:rawdatacomparedwithmodelestimates}, which follows @Alexander2019, shows the mean of the distribution and 95 per cent prediction intervals (as a reminder, prediction intervals tend to be wider than confidence intervals). In general, the raw estimate of the Australian Labor Party's share of first preferences in both surveys is reduced by the model. Figure \ref{fig:rawdatacomparedwithmodelestimates} is on the basis of electoral divisions, however it is also possible to do a similar exercise on the basis of, say, age-group, gender, or education.

@@

The classifications used by the Australian Electoral Commission are usually used only on actual results, and so we adjust the central classification in a way that is more explicit about uncertainty. Specifically: 'safe loss' means a two-party-preferred share of less than 0.4; 'fairly safe loss' means between 0.4 and 0.44; 'marginal loss' means between 0.44 and 0.47; 'too close to call' means between 0.47 and 0.53; 'marginal win' means between 0.53 and 0.56; 'fairly safe win' means between 0.56 and 0.6; and finally, 'safe win' means a two-party-preferred share greater than 0.6. Table \ref{tab:haveagoyougetago} shows this disaggregation. Using this classification,  estimates based on the Life in Australia dataset tend to be more accurate than those based on the Smartvote Australia dataset.

@@

@@

In our model, we account for preferences in a way that propagates uncertainty by applying a model that relates first-preferences shares to two-party-preferred shares which has been trained on the past four elections. This results in increased variance, compared with a deterministic approach.

@@

<!-- Possibly no need to include this for job app

@@

Finally, in terms of communication, we have provided some variations designed to ensure that the probability nature of our model cannot be separated from central estimates. However, this is just a first step and there is much more that could be done, including survey work and field testing to examine how this could be done in a better way. -->

@@

<!--One weakness of our approach is that it is not electoral-division-specific, which means that there are nuances of specific electoral divisions that it misses. Better preferences model - take account of preferences deals, and also not lump all non-major parties in as other - separate them based on ideology so that there is at least some better flow.-->
